\Title

### 2022-10-25 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.

Let us highlight some innovations in the contributions presented in this paper. First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have observed to be less efficient. Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif. Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen. Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}. Another main contribution is to provide a model that is suitable for learning any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs. This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified several limits to our model which we will now detail. First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time. We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme. Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework. Such an architecture could provide promising speed gains for the calculations.  \note{pourrait parler du poster bernstein? de nadafian?}

>>>>

\todo{try also a quantization of the weights (binary?)}
%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.

Let us highlight some innovations in the contributions presented in this paper. First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we have observed to be less efficient. Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif. Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen. Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}. Another main contribution is to provide a model that is suitable for learning any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs. This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified several limits to our model which we will now detail. First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time. We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme. Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework. Such an architecture could provide promising speed gains for the calculations.  \note{pourrait parler du poster bernstein? de nadafian?}
